An elaborate data set on human gait and the effect of mechanical perturbations

### 2014-12-15 paper.tex
Made some formatting adjustments to listings for the 12pt font.

\begin{listing}
  \small
  \begin{minted}[gobble=4]{yaml}
    study:
       id: 1
       name: Gait Control Identification
       description: Perturb the subject during walking and running.
    subject:
       id: 8
       age: 20
       mass: 70.0
       mass-units: kilograms
       height: 1.572
       height-units: meters
       knee-width-left: 107.43
       knee-width-right: 107.41
       knee-width-units: millimeters
       ankle-width-left: 70.52
       ankle-width-right: 67.66
       ankle-width-units: millimeters
       gender: male
    trial:
       id: 58
       subject-id: 8
       datetime: 2014-03-28
       notes: >
           The subject did a somersault during this trial instead of following
           instructions to walk. Will have to use for another study.
       nominal-speed: 0.8
       nominal-speed-units: meters per second
       stationary-platform: True
       pitch: False
       sway: False
       hardware-settings:
           high-performance: True
       dflow-version: 3.16.1
       cortex-version: 3.1.1.1290
       marker-map:
           M1: LHEAD
           M2: THEAD
           M3: RHEAD
           M4: FHEAD
           M5: C7
       analog-channel-map:
           Channel1.Anlg: F1Y1
           Channel2.Anlg: F1Y2
           Channel3.Anlg: F1Y3
           Channel4.Anlg: F1X1
       events:
           A: Force Plate Zeroing
           B: Calibration Pose
           C: First Normal Walking
           D: Longitudinal Perturbation
           E: Second Normal Walking
           F: Unloaded End
       files:
           compensation: ../T057/mocap-057.txt
           mocap: mocap-058.txt
           record: record-058.txt
           meta: meta-058.yml
  \end{minted}
  \caption{A fictitious example of a YAML formatted meta data file. All of the
    possible keys in the data set are shown.}
  \label{lis:example-meta-data}
\end{listing}

@@

Once the data is cleaned there are two methods that allow you to extract the
cleaned data: either extract sections of the data bounded by the events
recorded in the \verb|record-xxx.txt| file or save the cleaned data to disk.
These operations are available as a command line application and as an
application programming interface (API) in Python. An example of the
\verb|DFlowData| API in use is provided in Listing~\ref{lis:dflowdata}.
%
\begin{listing}
  \begin{minted}[gobble=4]{python}
    >>> from gaitanalysis.motek import DFlowData
    >>> data = DFlowData('mocap-020.txt', 'record-020.txt', 'meta-020.yml')
    >>> mass = data.meta['subject']['mass']
    >>> data.clean_data()
    >>> event_df = dflow_data.extract_processed_data(
    ...     event='Longitudinal Perturbation')
  \end{minted}
  \cprotect\caption{Python interpreter session showing how one could load a
    trial into memory, extract the subject's mass from the meta data, run the
    data cleaning process, and finally extract a Pandas \verb|DataFrame|
    containing all of the time histories for a specific event in the trial.}
  \label{lis:dflowdata}
\end{listing}

The \verb|GaitData| class is then used to compute things such as gait events
(toe off and heel strike times), basic 2D kinematics and inverse dynamics, and
to store the data into a Pandas \verb|Panel| with each gait cycle on the item
axis at a specified sample rate. This object can also be serialized to disk in
HDF5 format. An example of using the Python API is shown in
Listing~\ref{lis:gaitdata}.
%
\begin{listing}
  \begin{minted}[gobble=4]{python}
    >>> from gaitanalysis.gait import GaitData
    >>> gdata = GaitData(event_df)
    >>> gdata.inverse_dynamics_2d(left_markers, right_markers, left_loads,
    ...                           right_loads, mass, 6.0)
    >>> gdata.grf_landmarks('Right Fy', 'Left Fy', threshhold=20.0)
    >>> gdata.split_at('right')
    >>> gdata.plot_gait_cycles('Left Hip Joint Torque', mean=True)
    >>> gdata.save('gait-data.h5')
  \end{minted}
  \cprotect\caption{Python interpreter session showing how one could use the
    \verb|GaitData| class to load in the result of \verb|DFlowData| and compute
    the inverse dynamics (joint angles and torques), identify the gait event
    (e.g. heel strikes), split the data with respect to the gait events in a
    Pandas \verb|Panel|, plot the mean and standard deviation of one time
    history with respect to the gait cycles, and save the data to disk.}
  \label{lis:gaitdata}
\end{listing}

>>>>

\begin{listing}
  \scriptsize
  \begin{minted}[gobble=4]{yaml}
    study:
       id: 1
       name: Gait Control Identification
       description: Perturb the subject during walking and running.
    subject:
       id: 8
       age: 20
       mass: 70.0
       mass-units: kilograms
       height: 1.572
       height-units: meters
       knee-width-left: 107.43
       knee-width-right: 107.41
       knee-width-units: millimeters
       ankle-width-left: 70.52
       ankle-width-right: 67.66
       ankle-width-units: millimeters
       gender: male
    trial:
       id: 58
       subject-id: 8
       datetime: 2014-03-28
       notes: >
           The subject did a somersault during this trial instead of following
           instructions to walk. Will have to use for another study.
       nominal-speed: 0.8
       nominal-speed-units: meters per second
       stationary-platform: True
       pitch: False
       sway: False
       hardware-settings:
           high-performance: True
       dflow-version: 3.16.1
       cortex-version: 3.1.1.1290
       marker-map:
           M1: LHEAD
           M2: THEAD
           M3: RHEAD
           M4: FHEAD
           M5: C7
       analog-channel-map:
           Channel1.Anlg: F1Y1
           Channel2.Anlg: F1Y2
           Channel3.Anlg: F1Y3
           Channel4.Anlg: F1X1
       events:
           A: Force Plate Zeroing
           B: Calibration Pose
           C: First Normal Walking
           D: Longitudinal Perturbation
           E: Second Normal Walking
           F: Unloaded End
       files:
           compensation: ../T057/mocap-057.txt
           mocap: mocap-058.txt
           record: record-058.txt
           meta: meta-058.yml
  \end{minted}
  \caption{A fictitious example of a YAML formatted meta data file. All of the
    possible keys in the data set are shown.}
  \label{lis:example-meta-data}
\end{listing}

@@

Once the data is cleaned there are two methods that allow you to extract the
cleaned data: either extract sections of the data bounded by the events
recorded in the \verb|record-xxx.txt| file or save the cleaned data to disk.
These operations are available as a command line application and as an
application programming interface (API) in Python. An example of the
\verb|DFlowData| API in use is provided in Listing~\ref{lis:dflowdata}.
%
\begin{listing}
  \begin{minted}[gobble=4]{python}
    >>> from gaitanalysis.motek import DFlowData
    >>> data = DFlowData('mocap-020.txt', 'record-020.txt',
    ...                  'meta-020.yml')
    >>> mass = data.meta['subject']['mass']
    >>> data.clean_data()
    >>> event_df = dflow_data.extract_processed_data(
    ...     event='Longitudinal Perturbation')
  \end{minted}
  \cprotect\caption{Python interpreter session showing how one could load a
    trial into memory, extract the subject's mass from the meta data, run the
    data cleaning process, and finally extract a Pandas \verb|DataFrame|
    containing all of the time histories for a specific event in the trial.}
  \label{lis:dflowdata}
\end{listing}

The \verb|GaitData| class is then used to compute things such as gait events
(toe off and heel strike times), basic 2D kinematics and inverse dynamics, and
to store the data into a Pandas \verb|Panel| with each gait cycle on the item
axis at a specified sample rate. This object can also be serialized to disk in
HDF5 format. An example of using the Python API is shown in
Listing~\ref{lis:gaitdata}.
%
\begin{listing}
  \begin{minted}[gobble=4]{python}
    >>> from gaitanalysis.gait import GaitData
    >>> gdata = GaitData(event_df)
    >>> gdata.inverse_dynamics_2d(left_markers, right_markers,
    ...                           left_loads, right_loads, mass, 6.0)
    >>> gdata.grf_landmarks('Right Fy', 'Left Fy', threshhold=20.0)
    >>> gdata.split_at('right')
    >>> gdata.plot_gait_cycles('Left Hip Joint Torque', mean=True)
    >>> gdata.save('gait-data.h5')
  \end{minted}
  \cprotect\caption{Python interpreter session showing how one could use the
    \verb|GaitData| class to load in the result of \verb|DFlowData| and compute
    the inverse dynamics (joint angles and torques), identify the gait event
    (e.g. heel strikes), split the data with respect to the gait events in a
    Pandas \verb|Panel|, plot the mean and standard deviation of one time
    history with respect to the gait cycles, and save the data to disk.}
  \label{lis:gaitdata}
\end{listing}
